---
title: 'Statistical report: Association between clinical trial participation and outcomes in patients with heart failure: observations from the ESC Heart Failure Long-Term Registry'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    dev: cairo_pdf
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{draftwatermark}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

hf3_lt_fu_data_soladis_jan19.sas7bdat (from the folder DATABASE_ANALYZED.ZIP). 

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart")
```

First patient in: `r min(edata$num_dmVisitdt)` and last patient in: `r max(edata$num_dmVisitdt)`. 

The median age (IQR) is `r edata %>% summarise(med = fn(median(num_age, na.rm = TRUE), dig = 1),
                                             q1 = fn(quantile(num_age, na.rm = TRUE, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(num_age, na.rm = TRUE, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r edata %>% count(num_dmgender) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(num_dmgender == "Female") %>%
  pull(perc)`% females.  

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

For patients with missing information on the date of hospitalisation or date of hospitalisation prior to inclusion, 
the time to hospitalisation was imputed with half the time to last
follow-up. Further, times to hospitalisation that were larger than time 
to death or last follow-up were set to death or last follow-up. 

Missing data for the covariates included in the models was imputed with multiple 
imputation using mice [@mice] for 10 datasets and 10 iterations. 
Variables included in the imputation model are the same as are 
included in the cox regression. TP was not included in the model.  

\clearpage 

## Baseline characteristics

```{r, child = "../src/tab1.Rmd"}

```

Note 

- for substances = "Other" no doses are shown (can't be converted)
- for patients taking more than one acei/arb/arni the max dose will be presented (not summeraized). 

\clearpage

## Assocation between trial participation and characteristics

Performed using logistic regression models on the analysis population (patients with follow-up data).

```{r, child = "../src/predictors.Rmd"}

```

\clearpage

## Outcome analysis

The outcome analysis is performed on the analysis population (patients with follow-up data).

The following outcomes are considered: 

- All cause mortality
- First HF hospitalisation

Time is from date of discharge, or if this is missing, date of admission and 
censored at death not defined as an event or follow-up.  

The outcomes are presented with cumulative incidence curves and Cox proportional hazards regressions were 
used to model the time to first event analysis. Adjustment was performed for variables indicated in Table \ref{tab:tab1surv} 
and the selection was based on clinical judgment. 
Due to non-proportional hazards location was stratified for in the model.

The median (q1-q3) follow-up for the long-term outcomes is 
`r edata %>% filter(survpop) %>% summarise(med = fn(median(outtime_death / 365.25 * 12), dig = 1),
                                             q1 = fn(quantile(outtime_death / 365.25 * 12, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(outtime_death / 365.25 * 12, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` months for a total of 
                                   `r edata %>% filter(survpop) %>% 
                                   summarise(sumpy = fn(sum(outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.


```{r, child = "../src/km.Rmd"}

```

```{r, child = "../src/outtab.Rmd"}

```

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]). Due to non-proportional hazards location was 
stratified for in the model. 

### Subgroups

Subgroup analysis are performed for selected variables. 
Performed using interaction effects on the patients in the analysis population. 

```{r, child = "../src/subout.Rmd"}

```

\clearpage

### Residual confounding

In order to assess how robust the association between trail participation and the outcomes are 
to residual confounding the E-value 
was calculated [@evalue]. The E-value is interpreted as the minimum Risk Ratio 
that an unmeasured confounder would need to have with both trail participation and 
the outcome, conditional on the variables currently adjusted 
for in the model, to fully eliminate the association between trail participation and 
outcome. See Table \ref{tab:evalueprint2} in order to compare the magnitude of the E-value 
within the context of the study [@evalue2] (Table \ref{tab:evalueprint2} only for internal use, do not present in paper). 

```{r, child = "../src/evalue.Rmd"}

```

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/esctrialparticipation. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
